---
title: "Task1"
author: "Sukenik Peter"
date: "21/10/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 Atomic types

Question description

```{r}
# code for solving task
#numeric
population <- 56000
area <- 73
avg_temp <- 10.5
bridges <- 5
#integer
museums<- 3
monuments<-15
schools <- 109
#logical
UNESCO <- FALSE
green_transport <- TRUE
#character
city_name <- "Trencin"
river_name <- "Vah"
famous_monument <- "Trenciansky hrad"
```

## Question 2 Vectors

Question description

```{r}
# code for solving task
#numeric vector
num_vector <- c(population,area, avg_temp, bridges, museums,monuments,schools)

#sequence vector
s_vector <- seq(from = 15, to = 100, by =5)

#logical vector
l_vector <- sample(c(TRUE,FALSE),size=10, replace = TRUE)
```

## Question 3 Matrix

Question description

```{r}
# code for solving task
x <- sample(-50:50, 9, replace = TRUE)
m <- matrix(x, nrow = 3, ncol = 3)
rownames(m) <- c("R1","R2","R3")
colnames(m) <- c("C1","C2","C3")

row_means <- rowMeans(m)
col_means <- colMeans(m)

```

## Question 4 List

Question description

```{r}
# code for solving task
T_list <- list(population = population,
               area=area,
               avg_temp=avg_temp,
               museums=museums,
               bridges=bridges,
               monuments=monuments,
               schools=schools,
               UNESCO=UNESCO,
               green_transport=green_transport,
               city_name=city_name,
               river_name=river_name,
               famous_monument=famous_monument,
               num_vector=num_vector<-c(area,avg_temp, bridges,museums,monuments,schools),
               s_vector=s_vector <- seq(from = 15, to = 100, by =5),
               l_vector=l_vector <- sample(c(TRUE,FALSE),size=10, replace = TRUE),
               m=m,
               T_descript = 'Trencin, miesto kde sa streta kultura s modernym umenim')

```


## Question 5 Factor

Question description

```{r}
# code for solving task
development_level <- factor(c('very low','low', 'medium','high','very high'), levels = c('very low','low', 'medium','high','very high'))

table(development_level)
```

## Question 6 Data frame

Question description

```{r}
# code for solving task
cities_data <- data.frame(
  cities = c('Trencin', 'Zilina', 'Kosice', 'Presov', 'Nitra', 'Bratislava'),
  population = c(56000, 83000, 238000, 89000, 78000, 475000),
  area = c(73, 80, 242, 70, 100, 367),
  unesco = c(FALSE, FALSE, FALSE, FALSE, FALSE, TRUE)
)

colnames(cities_data) <- c('City', 'Population', 'Area', 'UNESCO')

str(cities_data)
summary(cities_data)

cities_data$Density <- cities_data$Population / cities_data$Area

cities_data
```

##TASK 2: Subsetting & Missing Data

### 2.1. Querying Vectors

```{r}
# code for solving task
# 2.1.1. Sequence vector
# 3th element
third_el <- s_vector[3]
third_el

# 2 to 5
two_five <- s_vector[2:5]
two_five

#greater than 20
greater_20 <- s_vector[s_vector > 20]
greater_20

# 2.1.2. Logical vector
# true values
true_vl <- l_vector[l_vector]
true_vl

# index
true_pst <- which(l_vector)
true_pst
```

### 2.2 Querying Your "Universe" List
```{r}
# code for solving task
# Using the `$` operator
T_list$T_descript

# matrix [[]]
T_list[['m']]

# second element
T_list[["num_vector"]][2]

```

### 2.3. Investigating Your Data Frame
```{r}
# code for solving task
# first 3 rows
head(cities_data, 3)

# 4th row, 1st column
cities_data[4, 1]

# contains logical value
cities_data$UNESCO
```

### 2.4. Handling "Corrupted" Data (`NA`)
```{r}
# code for solving task
#copying data
df_corrupted <- cities_data
#NA
df_corrupted[2, "Area"] <- NA
df_corrupted[5, "Population"] <- NA
df_corrupted[6, "UNESCO"] <- NA
#count NA
sum(is.na(df_corrupted))
#count column NA 
colSums(is.na(df_corrupted))
#mean of population with NA
mean(df_corrupted$Population)
#mean of population without NA
mean(df_corrupted$Population, na.rm = TRUE)
#remove NA
df_clean <- na.omit(df_corrupted)
df_clean
```

### 2.5. Bonus Challenge which
```{r}
# code for solving task
#highest population
highest_pop_index <- which.max(cities_data$Population)
cities_data[highest_pop_index, ]
#lowest destiny
lowest_density_index <- which.min(cities_data$Density)
cities_data[lowest_density_index, ]
```

## TASK 3: Expanding Your Universe

### 3.1. The Quest: Find Your Data
```{r}
# code for solving task
# import from: 'https://www.kaggle.com/datasets/juanmah/world-cities'
library(readr)
worldcities <- read_csv("worldcities.csv")
View(worldcities)

head(worldcities)
str(worldcities)
summary(worldcities)
```

```{r}
# code for solving task

```

```{r}
# code for solving task

```